manuscript_figures

### 2024-07-05 scripts/data_cleaning.Rmd
updasted HPI and production plots

>>>>

# Rename columns in species_metadata_type to match those in QUAD_data
colnames(species_metadata_type) <- c("SP_CODE", "type")

# Perform a left join to update the type column in QUAD_data
QUAD_data <- QUAD_data %>%
  left_join(species_metadata_type, by = "SP_CODE", suffix = c("", ".new"))

# Update the original type column with the new type values where the original type is NA
QUAD_data$type <- ifelse(is.na(QUAD_data$type), QUAD_data$type.new, QUAD_data$type)

# Remove the temporary new type column using base R
QUAD_data$type.new <- NULL

### 2024-07-05 scripts/manuscript_figures.Rmd
updasted HPI and production plots

@@

# Calculate fish biomass 'W' using the length-weight relationship
PB_filtered$biomass <- with(PB, coef.a * (actual_size)^coef.b)

# Divide biomass by 80 m to get g/m2
#calculate biomass/m2
PB$biomass.g.m2 <- PB$biomass / 80

# Constants from the P/B ratio equation
coefficient <- 2.64
exponent <- -0.35

# Calculate P/B ratio, the predictive equation is P/B (per year) = 2.64*Wmat^-035 , where Wmat is weight-at-maturity (Randall and Minns 2000)
PB$P_B_ratio <- coefficient * PB$biomass.g.m2^exponent

# Calculate Habitat Productivity Index (HPI), habitat productivity index (HPI), a proposed measure of habitat suitability and productive capacity, can be calculated as the product of average fish biomass and the species P/B coefficient (HPI = B × P/B) (Randall and Minns 2000)
PB$HPI <- PB$biomass.g.m2 * PB$P_B_ratio

# Aggregate by Transect, Site, Year and calculate mean HPI
hpi_summary <- PB %>%
  group_by(TRANSECT, SITE, YEAR) %>%
  summarise(mean_HPI = mean(HPI, na.rm = TRUE))

# Make faceted line plot of HPI index over the years
hpi.plot < - ggplot(hpi_summary, aes(x = interaction(TRANSECT, SITE, sep = " - "), y = mean_HPI, fill = as.factor(YEAR))) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  labs(x = "Transect - Site", y = "Mean Habitat Productivity Index (HPI)", fill = "Year") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1));hpi.plot

@@

# Define custom shape values for more than 8 categories
custom_shape_values <- c(16, 17, 18, 19, 20, 21, 22, 23)

>>>>

# Rename kings beach and fort wetherill
PB$SITE[PB$SITE == "Kings Beach"] <- "King's Beach"
PB$SITE[PB$SITE == "Fort Weatherill"] <- "Fort Wetherill"

@@

# Calculate fish biomass 'W' using the length-weight relationship
PB_filtered$Wmat <- PB_filtered$coef.a * (PB_filtered$Lmat^PB_filtered$coef.b)

# Constants for the PB.ratio calculation
coefficient <- 2.64
exponent <- -0.35

# Calculate PB.ratio column
PB_filtered$PB.ratio <- coefficient * (PB_filtered$Wmat ^ exponent)

# Calculate fish biomass
# Calculate total weight for each fish species
PB_filtered$biomass <- PB_filtered$Wmat * PB_filtered$COUNT

# Calculate biomass per m²
PB_filtered$biomass_per_m2 <- PB_filtered$biomass / 80

# Calculate production for each species in each transect
production_data <- PB_filtered %>%
  group_by(TRANSECT, SP_CODE, YEAR) %>%
  summarise(
    Total_Biomass = sum(biomass_per_m2, na.rm = TRUE),  # Sum of biomass for the species in the transect
    Total_Production = sum(biomass_per_m2 * PB.ratio, na.rm = TRUE)  # Total production for the species in the transect
  )

# Determine the maximum and minimum values for Total_Production
max_prod <- max(production_data$Total_Production)
min_prod <- min(production_data$Total_Production)

# Plotting using ggplot2
production.plot <- ggplot(production_data, aes(x = YEAR, y = Total_Production, fill = SP_CODE)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Year",
       y = "Fish Production",
       fill = "Species") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels if needed
        axis.title.x = element_text(vjust = -0.5),  # Adjust x-axis title position
        axis.title.y = element_text(vjust = 1.5)) + # Adjust y-axis title position
  scale_y_continuous(limits = c(0, 650), expand = c(0, 0));production.plot  # Set y-axis limits and remove expansion

# Calculate Habitat Productivity Index (HPI), habitat productivity index (HPI), a proposed measure of habitat suitability and productive capacity, can be calculated as the product of average fish biomass and the species P/B coefficient (HPI = B × P/B) (Randall and Minns 2000)
# Calculate average biomass per SP_CODE
average_biomass_df <- PB_filtered %>%
  group_by(SP_CODE, PB.ratio, YEAR, TRANSECT, SITE) %>%
  summarise(average_biomass = mean(biomass_per_m2, na.rm = TRUE))

# Calculate Habitat Productivity Index (HPI)
average_biomass_df$HPI <- average_biomass_df$average_biomass * average_biomass_df$PB.ratio

# Aggregate by Transect, Site, Year and calculate mean HPI
hpi_summary <- average_biomass_df %>%
  group_by(SITE, YEAR) %>%
  summarise(mean_HPI = mean(HPI, na.rm = TRUE))

# Make faceted line plot of HPI index over the years as a area plot
hpi.plot <- ggplot(hpi_summary, aes(x = YEAR, y = mean_HPI, fill = SITE)) +
  geom_area(position = "stack") +
  labs(x = "Year",
       y = "Mean Habitat productivity index",
       fill = "Site") +
  theme_minimal(); hpi.plot

@@

# Define custom shape values for more than 8 categories
custom_shape_values <- c(0, 1, 2, 3, 4, 5, 6, 23)
